\title{Availability, reproducibility and reusability evaluation levels: guideline for authors}

\section{Introduction}

Public policies promoting open science aim to develop the circulation of knowledge across all levels of society. They seek to strengthen the international influence of academic research while promoting transparency, cumulative growth and reproducibility in research. By sharing research methods, materials and results, these policies encourage the wider dissemination of knowledge beyond the academic world to society, communities and the economic sector. These policies are implemented at the global (UNESCO), continental (European Union) and national (France) levels.

\subsection{What is the interest ?}

Adopting open science practices can enhance the quality, credibility, and reach of your research. Open science is a collaborative effort that welcomes everyone—regardless of role or experience—to participate in creating a more equitable and trustworthy research ecosystem (COS, Center for Open Science). 

These principles aim to democratize access to research, promote equitable resource distribution, foster accountability and trustworthiness, accelerate self-correction, and improve rigor and reproducibility.

Reproducibility—the ability of independent researchers to obtain the same (or similar) results when repeating an experiment or test—is one of the hallmarks of good science (Popper 1959). Reproducibility provides scientists with evidence that research results are objective and reliable and not due to bias or chance (Rooney et al 2016).

Moreover, for a researcher, being able to reproduce his own work few months later allows to save a great amount of time.

\subsection{Increasing of the number of citations ?}
An international research team from PLOS Open Research Solutions has conducted a comprehensive study covering an entire country and all disciplines. The authors, Giovanni Colavizza (Denmark), Lauren Cadwallader (United States) and Iain Hrynaszkiewicz (United States), chose to study France, a leading country in monitoring Open Science, and to use data from the French Open Science Monitor, that covers the entire country since 2018.
The study covers a considerable corpus of over 500,000 scientific articles. \textbf{It reveals that each open science practice seems to be associated with an increase of the number of citations of the articles concerned.}
See \url{https://www.ouvrirlascience.fr/for-the-first-time-on-a-scale-of-a-country-france-a-study-demonstrates-that-open-science-could-increase-the-chances-for-researchers-to-be-cited/}.
An other source (\url{https://www.replicabilitystamp.org/}) also stats that papers with the Replicability Stamp are cited 15$\%$ more than average one year after publication, and 33$\%$ more two years after publication. [Calculated based on Clarivate Web Of Science (WoS) citation counts in 2020].

\section{Data management}

\section{Code management}
We distinguish 3 different independent categories (some badges may be obtained independently):

\begin{itemize}
\item Availability: the software artefacts used in the publication adhere to the Open Science standard. Reader can find the code in an open repository, with mandatory information such as authors and licence files. 

\item Reproducibility: readers can reproduce the main results of the publication. The authors have provided all the necessary elements to enable this.
  
\item Reusability: The software artefacts can be reused for goals or experiments other than those described in the article. The authors have provided readers with all the elements needed to modify the code, add new functionalities, and ensure cumulative science. 
\end{itemize}

In the following section, for each of them, a finer granularity is defined to indicate what can be done to get more badges.


\subsection{Availability of the code}

As described before, to respect the standard of open science, the code used in the publication has to be available for reviewers and readers. We details here the different levels of availability of the code:\\

\begin{tabular}{|p{10cm}|p{1.5cm}|}
%\begin{tabular}{|c|c|}
  \hline
  \cellcolor{bl1}  \cellcolor{bl1} Description & \cellcolor{bl1} Stamp \\
\hline
code available on a public forge + licence, authors, contributors files, commit number and branch corresponding to the publication & \textcolor{green1}{\ding{78}} \\
\hline
+ code available on Zenodo & \textcolor{green1}{\ding{78}\ding{78}} \\
\hline
+ Software Heritage Identifier (SWHID) &  \textcolor{green1}{\ding{78}\ding{78}\ding{78}}\\

\hline
+ HAL deposit & \textcolor{green1}{\ding{78}\ding{78}\ding{78}\ding{78}} \\ 
\hline
\end{tabular}
~\\

More details :
\begin{itemize}
\item Level 1: \\
without a licence file, the code cannot be used by other people. The files of the authors and contributors are needed for patrimonial and author's rights. Finally, providing the commit and the exact branch enables further modifications to be made in the repository by author, without losing the state corresponding to the article.   

\item Level 2: \\
Zenodo provides a DOI and enables to fix a precise state of the software.
  
\item Level 3: \\
Hosting your source code on the Software Heritage platform enables a perennial archive and a very precise commit identification (SWHID) (for example, in the article, you can cite a specifi script with a specific identifier). 
\item Level 4: \\
  HAL deposit enables to verify that all the needed files are present and in the right format (curation).\\
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Reproducibility}

\begin{tabular}{|p{10cm}|p{1.6cm}|p{2.2cm}|}
\hline
\cellcolor{bl1} Description & \cellcolor{bl1} Stamp & \cellcolor{bl1} Requirement \\
\hline
re-runable code, description of the workflow and test case(s) &  \textcolor{yell1}{\ding{115}} & \textcolor{green1}{\ding{78}}\\ 
\hline
+ workflow manager for automatic execution  &  \textcolor{yell1}{\ding{115}\ding{115}} & \textcolor{green1}{\ding{78}}\\ 
\hline
+ reproducible software environment & \textcolor{yell1}{\ding{115}\ding{115}\ding{115}} & \textcolor{green1}{\ding{78}} \\
\hline
+ regression, test cases and documentation (in CI ?) & \textcolor{yell1}{\ding{115}\ding{115}\ding{115}\ding{115}} &  \textcolor{green1}{\ding{78}} \\
 \hline
\end{tabular}
~\\

More details :
\begin{itemize}
\item Level 1: \\
Supposing the code is available, the reader can re-run it understanding which component has to be used and in which order. She/he can also verify through a test that he obtains correct results.    

\item Level 2: \\
A workflow manager solution is provided: the reader has just to launch the procedure in order to run a test case. As workflow manager tool, we could cite snakemake and nextflow for example. 
  
\item Level 3: \\
The software environement is entirely described: the reader can deploy it in his own computer (automatically or at leat manually) with the correct version number of dependances. \textcolor{red}{Various tools simplify this step such as nix/guix, cmake for compilation, etc.}

\item Level 4: \\
The repository contains regression tests, test cases and automatic documentation \textcolor{red}{in a CI}: any modification is tested and does not modify the former behaviour of the code in specific configurations. Moreover, a new documentation is automatically generated.   
\end{itemize}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Reusability for further projects}


\begin{tabular}{|p{12cm}|p{1.6cm}|p{2.2cm}|}
\hline
\hline 
\cellcolor{bl1} Description & \cellcolor{bl1} Stamp & \cellcolor{bl1} Requirement \\
\hline
All functions documentation inside the code & \textcolor{purple1}{\ding{112}} &  \textcolor{yell1}{\ding{115}} \\
\hline
+ Description of each element in a Readme + workflow description for execution &  \textcolor{purple1}{\ding{112}\ding{112}} & \textcolor{yell1}{\ding{115}} \\
\hline
+ User documentation with examples &  \textcolor{purple1}{\ding{112}\ding{112}\ding{112}}&\textcolor{yell1}{\ding{115}} \\
\hline
+ Developper documentation (``How to contribute ?'' guide) &  \textcolor{purple1}{\ding{112}\ding{112}\ding{112}\ding{112}}&\textcolor{yell1}{\ding{115}} \\
\hline
\end{tabular}
~\\

More details :
\begin{itemize}
\item Level 1: \\
Each module of the code is highly documented with useful tools as sphynx or doxygen for example  

\item Level 2: \\
The role of each piece of code or script is documented in a Readme file. Each step of execution is described in order to facilitate the reusability.   
\item Level 3: \\
The ``user'' documentation is complete with simple test cases, examples, links with bibliography and theoretical elements. Numerical methods are described, as is the rationale behind their selection.   
\item Level 4: \\
The repository contains a guide for future developpers, explaining the choices made during the software development and the way to modify it. 
\end{itemize}





%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
